A multi-scalar story of the diffusion of a new technology: the web

### 2023-12-14 abstracts_notes/abstract_GfR_2024/Tranos_GfR_2024.tex
local changes

\documentclass[11pt]{article}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage[demo]{graphicx}
%\usepackage{subfig}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%\usepackage[square,numbers]{natbib}
\usepackage[round]{natbib} % remove number, use ()
\bibliographystyle{abbrvnat}

@@

% remove date
%\date{}

% ========== Edit your name here
\author[]{Professor Emmanouil Tranos}
\affil[]{University of Bristol and The Alan Turing Institute}

@@

\begin{center}
\href{mailto:e.tranos@bristol.ac.uk}{e.tranos@bristol.ac.uk},
\href{https://etranos.info/}{etranos.info},
\href{https://twitter.com/EmmanouilTranos}{@EmmanouilTranos} \&
\href{https://www.linkedin.com/in/emmanouil-tranos-42947812a/}{LinkedIn}
\end{center}

@@

% general, summary
This paper maps the participation in the digital economy and its evolution in the UK over space and time. Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects. Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour. Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK. Instead of metrics capturing the passive engagement with digital technologies, for instance internet subscription metrics, this paper targets the active engagement with the digital as reflected in website creation.

Importantly, the data and geolocation strategy allow to capture these processes at small spatial scales. This level of granularity differentiate this paper with previous approaches in the literature, which were only able to capture technological adoption at more coarse level. Thus, this paper tests how well established theoretical approaches regarding the diffusion of technologies are still applicable when the focus is on local scales. Although we know that spatial contagion and urban hierarchies are key drivers of technological diffusion, such theoretical concepts have not been tested at small geographical scales. Apart from an empirical interest, understanding such processes at small scales can support designing strategies for the adoption of new technologies and the development of relevant infrastructure.

The paper begins with simple analytical tools which reveal the scale-free character of the 'S curve' that defines the diffusion of new technologies and innovations. Indeed, such an 'S curve' describes the adoption of web technologies even at the level of local authorities in the UK. Then the paper continues with other exploratory approaches, which signal the spatial dependence of the diffusion on web technologies. In the last section, the paper employs state of the art machine learning techniques to predict the diffusion of web technologies over space and, importantly, highlight how different or not UK regions have been in terms of the adoption of web technologies.

Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.

\begin{figure}%
    \centering
    \subfloat[\centering label 1]{{\includegraphics[width=5cm]{./lad_E07000031.png} }}%
    \qquad
    \subfloat[\centering label 2]{{\includegraphics[width=5cm]{./lad_E07000073} }}%
    \caption{2 Figures side by side}%
    \label{fig:example}%
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}{.3\textwidth}
		\includegraphics[width=\textwidth]{lad_E07000031.png}
		\caption{Atom 1}
	\end{subfigure}
%%%%%%%%%%%%%%
	\begin{subfigure}{.3\textwidth}
		\includegraphics[width=\textwidth]{lad_E07000031.png}
		\caption{Atom 2}
	\end{subfigure}
%%%%%%%%%%%%%%
	\begin{subfigure}{.3\textwidth}
		\includegraphics[width=\textwidth]{lad_E07000031.png}
		\caption{Atom 3}
	\end{subfigure}
	\caption{Atoms are fun!}
\end{figure}

>>>>

\documentclass[11pt]{article}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{ {./} }

\usepackage[square,numbers]{natbib}
\bibliographystyle{abbrvnat}

@@

% ========== Edit your name here
\author[1]{Prof Emmanouil Tranos}
\affil[1]{School of Geographical Sciences, University of Bristol and The Alan Turing Institute}

@@

\begin{center}
\href{mailto:e.tranos@bristol.ac.uk}{e.tranos@bristol.ac.uk},
\href{https://etranos.info/}{etranos.info}
%\href{https://twitter.com/EmmanouilTranos}{@EmmanouilTranos} \&
%\href{https://www.linkedin.com/in/emmanouil-tranos-42947812a/}{LinkedIn}
\end{center}

@@

This paper maps and models the participation in the digital economy and its evolution in the UK over space and time. Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects. Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour. Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK. Instead of using metrics capturing the passive engagement with digital technologies, for instance internet subscription metrics, this paper targets the active engagement with the digital as reflected in website creation.

Importantly, the data and geolocation strategy allow to capture these processes at small spatial scales. This level of granularity differentiate this paper with previous approaches in the literature, which were only able to capture technological adoption at more coarse level. Thus, this paper tests how well established theoretical approaches regarding the diffusion of new technologies are still applicable when the focus is on local scales. Although we know that spatial contagion and urban hierarchies are key drivers of technological diffusion, such theoretical concepts have not been tested at small geographical scales. Apart from an empirical interest, understanding such processes at small scales can support designing strategies for the adoption of new technologies and the development of relevant infrastructure.

The paper begins with simple analytical tools which reveal the scale-free character of the 'S curve' that defines the diffusion of new technologies and innovations. Indeed, such an 'S curve' describes the adoption of web technologies across scales and even at the level of local authorities in the UK (Figure \ref{fig:s}). Then, the paper continues with exploratory approaches, which signal the spatial dependence of the diffusion on web technologies. All the above provide useful insights for the last section, when the paper employs state of the art machine learning techniques to predict the diffusion of web technologies over space. Specifically, Random Forest models are trained at two different scales: local authorities (c. 400 in the UK) and census output areas (c. 230,000). The models are firstly trained and tested in all data using space and time sensitive \emph{k}-folds. This modelling exercise highlights the importance of the key spatial predictors as described in the literature in predicting technological diffusion. Then, these model are trained for each UK region separately and tested iteratively to all other regions in order obtain an understanding of how differently these spatial drivers affect the diffusion of web technologies across different places (Figures \ref{fig:la} and \ref{fig:oa}). In other words, this modelling strategy reveals how different or not UK regions have been in terms of the adoption of web technologies across scales. The granularity of the data used here allows to expose spatial differences that otherwise would have gone unnoticed. The results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{lad_E09000001}
    \caption{Example of an S-curve for a local authority}
    \label{fig:s}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{heatmap_LA}
    \caption{Regional similarities based on cross-region training/predictions at the Local Authority level}
    \label{fig:la}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{heatmap_OA}
    \caption{Regional similarities based on cross-region training/predictions at the Local Authority level}
    \label{fig:oa}
\end{figure}

### 2023-12-14 presentations/web_diffusion.Rmd
local changes

---
title: 'Digital economy in the UK: an evolutionary story'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: bibliography.bib
biblio-style: "apalike"
css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

@@

::: {style="margin-left: 50px !important; text-align:left !important"}
-   Introduction
-   Web data
-   Spatial analysis
-   S-shaped diffusion curves
-   Rank dynamics
-   Conclusions <br> <br>
    <!-- [etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/) -->
:::

@@

## Aims

-   Map the **active** engagement with the *digital*
-   Over time, **early stages** of the internet
-   **Granular** and **multi-scale** spatial perspective

## Aims

@@

-   Understand how the adoption of new technologies evolves

@@

@@

-   Archived web data

@@

-   Geolocate:

    - to a unique postcode

    - to multiple postcodes

-------------------------------------------

@@

@@

![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)

@@

![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)

# Web diffusion

## Mapping website density

```{r heatmaps, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

---

-   Websites in an Output Area $i$

:::

-   Nearest retail centre

:::

- Spatial dependency
	- Relatively small and constant over time / scales
	- London hot spot early on
	- At local scale, consistent hotspots over time
	- Granular analysis reveals other hotspots

- Distance effect
	- Urban gravitation increases over time and then drops
	- Granular analysis: gravitation
	- Lost explanatory power over time

> **Hierarchy effect**: from main centers to secondary ones -- central
places

@@

```{r UK S, eval=T}
#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"

path.image <- paste0(path, "/outputs/s_uk.png")
knitr::include_graphics(path.image)
```

@@

```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
knitr::include_graphics(path.image)
```

@@

- Spatial heterogeneity
- Not clear pattern

@@

- Adoption heterogeneity 
- Perceptions of risk and economic returns from new technologies
- Early adopters vs. laggards, leapfrogging

@@

@@

----

-   **Geography matters:** spatial dependency, urban gravitation

@@

-   Well-established theoretical approaches of diffusion survive even at
    a granular level

-   Stability and volatility: leapfrogging, early adopters dropping, but also maintaining

>>>>

---
title: 'A multi-scalar story of the diffusion of a new technology: the web'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: bibliography.bib
biblio-style: "apalike"
css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

@@

::: {style="margin-left: 50px !important; text-align:left !important"}
-   Introduction

-   Web data

-   Methods

-   Results

    -   S-shaped diffusion curves
    -   Rank dynamics
    -   Spatial analysis
    -   Modelling

-   Conclusions <br> <br>
    <!-- [etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/) -->
:::

@@

## Aim

@@

@@

## Diffusion of a new *digital* technology

-   Diffusion of an intangible/digital technology

-   Map the **active** engagement with the *digital*

-   Over time, **early stages** of the internet

-   **Granular** and **multi-scale** spatial perspective

@@

-   Data from the Internet Archive, the oldest web archive

@@

-   Geolocation: postcode references in the text

-   Timestamp: archival year

-   Counts

@@

-   Data from a Web Archive -- The Internet Archive

-   Geolocation: postcode references in the text

-   Timestamp: archival year

-   Counts

@@

![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)

@@

![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)

# Methods

## Reminder: diffusion mechanisms

-   **S-shaped** pattern in the cumulative level of adoption

-   A **hierarchy effect**: from main centres to secondary ones

-   A **neighborhood effect**: first "hitting" nearby locations

## Methods

-   Model cumulative adoption

-   Descriptive statistics, ESDA & density regressions

-   Modelling framework

-   Two scales:

    -   websites per firm in a Local Authority
    -   websites in an Output Area

@@

```{r UK S, eval=T, out.width = "60%", out.height="60%"}
#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"

path.image <- paste0(path, "/outputs/s_uk_per_firm.png")
knitr::include_graphics(path.image)
```

@@

```{r Liverpool, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
knitr::include_graphics(path.image)
```

@@

-   Spatial heterogeneity

-   Not a clear, easy to explain pattern

@@

-   Adoption heterogeneity

-   Different perceptions of risk and economic returns from new
    technologies

-   Early adopters vs. laggards, leapfrogging

@@

## Rank dynamics and diffusion speed

-   Spatial heterogeneity

-   Expected volatility

# Spatial mechanisms

```{r, eval=FALSE, heatmaps, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

-   Websites in an Output Area $i$
:::

-   Nearest retail centre
:::

-   Spatial dependency
    -   Relatively small, consistent over time / scales
    -   London hot spot early on
    -   At local scale, consistent hotspots over time
    -   Granular analysis reveals other hotspots
-   Distance effect
    -   Urban gravitation increases over time and then drops
    -   Granular analysis: gravitation
    -   Lost explanatory power over time

> **Hierarchy effect**: from main centers to secondary ones -- central
> places

# Putting all of these together

## Modelling framework

-   Random forests to predict $Website\,Density_{i,t}$

-   4 sets of models:

    -   All the data ⟹ *variable importance*
    -   Train on one region, test on the rest ⟹ *spatial differences and
        similarities* of diffusion mechanisms
    -   For Local Authorities and Output Areas

-   Space-time sensitive 10-fold CV (`CAST`)

## Models trained on all data
<br>

|                   | RMSE  | RSquared | MAE   |
|-------------------|-------|----------|-------|
| Local Authorities | 0.022 | 0.908    | 0.012 |
| Output Areas      | 3.613 | 0.627    | 0.533 |

## Variable importance for Local Authorities

```{r varimpLA, eval = T, out.height= "58%", out.width="58%", message=FALSE}

path.image <- paste0(path, "/outputs/rf/varimp_LA.png")
knitr::include_graphics(path.image)
```

## Variable importance for Output Areas

```{r varimpOA, eval = T, out.height= "58%", out.width="58%", message=FALSE}
#,  fig.height=5, fig.width=5,message=FALSE

path.image <- paste0(path, "/outputs/rf/varimp_OA.png")
knitr::include_graphics(path.image)
```

## Regional similarities for Local Authorities

```{r similarity_LA, eval = T, message=FALSE}
#fig.height=5, fig.width=5 
#, out.height= "60%", out.width="60%"
path.image <- paste0(path, "/outputs/rf/heatmap_LA.png")
knitr::include_graphics(path.image)
```

## Regional similarities for Output Areas

```{r similarity_OA, eval = T, out.height= "70%", out.width="70%", message=FALSE}

path.image <- paste0(path, "/outputs/rf/heatmap_OA.png")
knitr::include_graphics(path.image)
```

@@

-   Established technological diffusion drivers still apply

    -   for a digital technology
    -   at local scales

-   Geography matters: spatial dependency, urban gravitation

@@

-   Stability and volatility: leapfrogging, early adopters dropping, but
    also stable positions

-   Spatially consistent mechanisms at local scale

-   Heterogeneity increases with resolution
